---
title: "How To Choose A Dog Based On Your Lifestyle"
author: "Leah Reichert"
date: " 11 May 2023"
output: rmdformats::material
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      warning = FALSE,
                      message = FALSE,
                      fig.path = "../Output/")
```



# Why I chose These Datasets 
- I wanted to help my friend narrow down what dog breed to get.



## **Load the Libraries** ## 
```{r}
library(plotly)
library(here)
library(tidyverse)
library(magick)
library(rmdformats)

```

## **Read in the Data** ## 
```{r}
Breed_Traits <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-02-01/breed_traits.csv')

Breed_Traits <- read_csv(here("Data","Breed_Traits.csv")) # Load first data set

head(Breed_Traits)


AKC_Breed_Info <- read_csv(here("Data","AKC_Breed_Info.csv")) # Load second data set 

head(AKC_Breed_Info)


```
# Breed Groups vs. Traits

## **Pivot Data Long** ## 
```{r}
long <- Breed_Traits %>%
 pivot_longer(cols = c(`Affectionate With Family`,`Good With Young Children`, `Good With Other Dogs`, `Shedding Level`, `Coat Grooming Frequency`, `Drooling Level`, `Openness To Strangers`, `Playfulness Level`, `Watchdog/Protective Nature`, `Adaptability Level`, `Trainability Level`, `Energy Level`, `Barking Level`, `Mental Stimulation Needs`), # variables in the new column
              names_to = 'Traits', # new name of the column
              values_to = 'Values') # name of the column 
head(long)
```

## **Create Bar Plot: Character traits measured by Breed Group** ## 
```{r, fig.width= 11, "Breed Groups vs. Traits"}
Traits_long <- long %>% # graph with this data 
  plot_ly(x = ~ `Breed Group`, 
          y = ~ Values,
          color = ~ Traits,
          type = "bar", # bar graph 
          marker = list(color = rainbow(nrow(long)))) %>% # bar graph is colored by rainbow
  layout(title = 'Breed vs. Traits', # add a title 
         xanchor = 'center', # center the title 
          yanchor = 'top',
         font = list(color = "darkorchid"), # change title color 
       plot_bgcolor = "white", # made background white
       paper_bgcolor = "lavender", # paper background color
       xaxis = list(title = 'Breed Group', # label x axis
                    tickangle = -45, # angle the x axis 
                    color = "darkorchid", # color x axis
                      size = 15),# make the x axis smaller 
       yaxis = list(title = 'Rank', # label y axis 
                    color = "darkorchid"), # color of y axis title 
       legend = list(title = list(text = '<b> Traits </b>', # rename legend 
                font = list(color = "darkorchid"), # change legend color 
                                  xanchor = 'center', # center the legend
                                  yanchor = 'top')))



Traits_long
```
##### **Trait Preference**  
1. Affectionate With Family    
2. Good With Other Dogs    
3. Adaptability Level  
4. Trainability Level  


# Narrow Down By Coat Type and Length

## **Filter Data To Visualize Coat Length** ##

##### **Coat Type Preference**  
- Smooth  
- Double  

##### **Coat Length Preference**  
- Short  

```{r}
coat_filtered <- Breed_Traits %>%
  filter(`Breed Group` %in% c("Hound","Sporting")) %>% # filter out breed group
  filter(`Coat Type` %in% c("Smooth","Double")) %>%# filter out coat preference
  filter(`Affectionate With Family` >= 3)%>% # filter following traits
  filter(`Good With Other Dogs` >= 3) %>%
  filter(`Trainability Level` >=3) %>%
  filter(`Adaptability Level` >=3) %>%
  filter(`Coat Length` == "Short") # filter by coat length 
  
  
head(coat_filtered)
```

## **Table: Coat Type for Certain Breeds** ##
```{r, fig.width = 11, fig.height= 15, "Coat Type/Length in Hounds and Sporting Groups"}
table <- coat_filtered %>%
plot_ly(
  type = 'table', # type of graph 
  columnwidth = c(80), # column width
  header = list(
    values = c('<b>Breed</b>','<b>Affectionate With Family</b>','<b>Good With Other Dogs</b>','<b>Adaptability Level</b>','<b>Trainability Level</b>','<b>Coat Type</b>','<b>Coat Length</b>'), # column names
    line = list(color = '#506784'), # line color
    fill = list(color = '#119DFF'),# column name color
    align = c('left','center'), # column names centered
    font = list(color = 'white', size = 12) #column name title color
  ),
  cells = list(
    values = rbind(
      c('Retrievers (Labrador)', 'Beagles', 'Pointers (German Shorthaired)', 'Dachshunds', 'Brittanys','Vizslas','Basset Hounds','Weimaraners','Rhodesian Ridgebacks','Bloodhounds','Whippets','Pointers','Black and Tan Coonhounds
','Bluetick Coonhounds','Greyhounds','Salukis','Redbone Coonhounds','Treeing Walker Coonhounds','Cirnechi dell’Etna','Pharaoh Hounds','American English Coonhounds','Sloughis','Harriers','American Foxhounds','English Foxhounds'), # row titles
      c(5,3,5,5,3,5,3,5,5,4,5,5,4,3,4,5,5,5,4,5,3,4,5,3,5),
      c(3,3,3,3,3,4,4,4,4,4,4,5,5,5,5,5,5,5,5,5,5,5,5,5,5),
      c(4,4,3,3,3,4,4,3,5,4,4,5,4,3,3,4,4,3,4,4,3,3,4,3,4),
      c(5,4,4,3,3,5,4,5,5,3,4,5,3,3,3,5,3,4,3,5,4,3,4,3,4), # input values
      c("Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Double","Smooth","Smooth","Smooth","Double","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Smooth","Double","Smooth","Double"), # input coat type 
      c("Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short","Short")), # input coat length
    line = list(color = '#506784'), # row line color
    fill = list(color = c('#25FEFD', 'white')), # row fill color 
    align = c('left', 'center'), # row title alignment 
    font = list(color = c('#506784'), size = 12) # row text color and size
    ))

table
```

# Narrow Down By Weight

## **Filter Data To Visualize Weight** ##  

##### **Weight Ranger Preference**
- 45-100lbs  
```{r}
weight_f <- AKC_Breed_Info %>%
  drop_na() %>%
  filter(Breed %in% c("Whippet","Vizsla","Rhodesian Ridgeback","Labrador Retriever","German Shorthaired Pointer","Pharaoh Hound","Harrier","English Foxhound","Black And Tan Coonhound","Basset Hound","Beagle","Dachshund","Brittany","Weimarener","Bloodhound","Whippet","Basenji","Pointer","Greyhound","Redbone Coonhound")) %>% # filter data to be universal 
  filter(weight_low_lbs >= 45) %>% # look at weights greater than 45 lbs
  filter(weight_high_lbs <= 100) # weight less than 100 lbs
  
  
head(weight_f)
```

## **Scatter Plot: Weight for Certain Breeds** ##
```{r, fig.width = 11, "Breeds vs. Weight"}
weight <- weight_f %>% # plot filtered data 
  plot_ly(x = ~ weight_low_lbs,
          y = ~weight_high_lbs,
          type = "scatter", # type of data
          mode = "markers",
          color = ~Breed, # color by
          colors = "Set1", # color pattern
          size = 20) %>% # size of the dot
layout(title = 'Breed vs. Weight', # title of plot
         font = list(color = "black"), # title color 
       plot_bgcolor = "white",# plot color 
       paper_bgcolor = "thistle", # background color 
       xaxis = list(title = 'Low Weight Range', # x axis title 
                    tickangle = -45, # angle x axis 
                    color = "black"), # color x axis
       yaxis = list(title = 'High Weight Range', # y axis title 
                    color = "black"), # y axis color 
       legend = list(title = list(text = '<b> Breed </b>', # legend title 
                                  font = list(color = "black"), # legend color 
                                  xanchor = 'center', # center title 
                                  yanchor = 'top')))

weight


```
# Narrow Down By Height

## **Filter Data Based on Height** ##

##### **Height Preference Range**  
- 21-26 inches  
```{r}
height_f <- weight_f %>%
  filter(height_low_inches >= 21) %>% # filter by 21 inches and greataer
  filter(height_high_inches <= 26) # filter by less than or equal to 26 inches 

head(height_f)
  
```

## **Scatter Plot: Height for Certain Breeds** ##
```{r, fig.width = 11, "Breed vs. Height" }
height <- height_f%>% # plot filtered data 
  plot_ly(x = ~ height_low_inches,
          y = ~height_high_inches,
          type = "scatter", # type of data
          mode = "markers",
          color = ~Breed, # color by
          colors = "Set1", # color pattern
          size = 22)%>% # size of the dot 
layout(title = 'Breed vs. Height', # title of plot
         font = list(color = "black"), # title color 
       plot_bgcolor = "white",# plot color 
       paper_bgcolor = "thistle", # background color 
       xaxis = list(title = 'Low Height Range', # x axis title 
                    tickangle = -45, # angle x axis 
                    color = "black"), # color x axis
       yaxis = list(title = 'High Height Range', # y axis title 
                    color = "black"), # y axis color 
       legend = list(title = list(text = '<b> Breed </b>', # legend title 
                                  font = list(color = "black"), # legend color 
                                  xanchor = 'center', # center title 
                                  yanchor = 'top')))
height
```
# One Last Comparison 

## **Filter Data For 4 Categories** ##

```{r}
trait_filter <- long %>%
 # mutate(Breed = trimws(Breed)) %>%
  filter(Breed %in% c("English Foxhounds" ,"Retrievers (Labrador)","Pharaoh Hounds" ,"Vizslas")) %>% # look at these breeds
  filter(Traits %in% c("Trainability Level","Adaptability Level","Good With Other Dogs","Affectionate With Family")) # look at these traits 

head(trait_filter)

```

## **Bar Plot: Final Filtered Breeds by Adaptability, Trainability, Good With Other Dogs, and Affectionate With Family** ##
```{r, fig.width = 11, "Final Breed vs. Trait"}
final <- trait_filter %>% # data used
  plot_ly(x = ~ Traits,
          y = ~Values,
          type = "bar", # type of graph 
          mode = "markers",
          color = ~Breed, # color by
          colors = "Paired") %>% # color pattern
  layout(title = 'Final Breed vs. Preferred Traits', # plot title
         font = list(color = "darkred"), # title color 
       plot_bgcolor = "bisque", # color of plot
       paper_bgcolor = "bisque", # color of background
       xaxis = list(title = 'Traits', # x axis title 
                    tickangle = -45, # angle x axis 
                    color = "darkred"), # x axis color
       yaxis = list(title = 'Rank',# y axis title
                    color = "darkred"), # y axis color
       legend = list(title = list(text = '<b> Breed </b>', # legend title
              font = list(color = "darkred"), # legend color
                      xanchor = 'center', # title centered 
                      yanchor = 'top')))


final


```
# Image of The Winner

## **Image of The Winner** ##
```{r, Winner, fig.width = 11}
Lab <- image_read("https://patchpuppy.com/wp-content/uploads/2022/05/AreEnglishLabsGoodDogs.jpg") # read in image

image_border(Lab, color = "chocolate4", 
             geometry = "26x28") %>% # Add a border
image_annotate("Labrador Retriever", font = 'Palatino',
               size = 50, 
               gravity = "southwest",
               color = "white", 
               boxcolor = "chocolate4") # add text box 

```

# References 

## **References** ##
> https://www.akc.org/expert-advice/dog-breeds/what-dog-is-right-for-me/
> https://www.masterclass.com/articles/dog-breed-guide









